
## Setting up the data
```{r}
### Install package
pacman::p_load(
    here, tidyverse, sva, bladderbatch, pamr, limma
    )
### Load data
data(bladderdata)

### Get the expression data
edata <- exprs(bladderEset)
head(edata)

### Get the variable data
pheno <- pData(bladderEset)
head(pheno)
```
##

## Adjusting batch effect using Linerar model
```{r}
mod <- model.matrix(~as.factor(cancer) + as.factor(batch), data = pheno)
fit <- lm.fit(mod, t(edata))
hist(fit$coefficients[2, ], col = 4, breaks = 100)
table(pheno$cancer, pheno$batch)
```

## Adjusting for batch effect with {Combat}

```{r}
batch <- pheno$batch
modcombat <- model.matrix(~1, data = pheno)
modcancer <- model.matrix(~cancer, data = pheno)
combat_edata <- ComBat(
    dat = edata, 
    batch = batch,
    mod = modcombat, 
    par.prior =  TRUE,
    prior.plots = FALSE
    )
combat_fit <- lm.fit(modcancer, t(combat_edata))
hist(combat_fit$coefficients[2, ], col = 4, breaks = 100)
```

## Comparing the combat and linear adjustment
We can compare the estimated coefficients from Combat and linear adjustment by looking at the right coefficients for each model.


```{r}
plot(
    fit$coefficients[2, ], combat_fit$coefficients[2, ],
    col = 4, xlab = "Linear Model", ylab = "Combat",
    xlim = c(-5, 5), ylim = c(-5, 5)
)
abline(c(0, 1), col = 2, lwd = 3)
dev.off()
```

## Adjusting for batch effects with {sva}


```{r}
mod <- model.matrix(~cancer, data = pheno)
mod0 <- model.matrix(~1, data = pheno)
sva1 <- sva(edata, mod, mod0, n.sv = 2)
summary(lm(sva1$sv ~ pheno$batch))

boxplot(sva1$sv[, 2] ~ pheno$batch)
points(sva1$sv[, 2] ~ jitter(as.numeric(pheno$batch)), col = as.numeric(pheno$batch))

### Add the surrogate variables to model matrix and perfom the model fit
modsv  <-  cbind(mod, sva1$sv)
fitsv <- lm.fit(modsv, t(edata))

### Compare the fit from surrogate variable analysis to the other two
par(mfrow=c(1,2))
plot(fitsv$coefficients[2,],combat_fit$coefficients[2,],col=2,
      xlab="SVA",ylab="Combat",xlim=c(-5,5),ylim=c(-5,5))
abline(c(0,1),col=1,lwd=3)
plot(fitsv$coefficients[2,], fit$coefficients[2,],col=2,
      xlab="SVA",ylab="linear model",xlim=c(-5,5),ylim=c(-5,5))
abline(c(0,1),col=1,lwd=3)
```
## Principal components for population structure


